---
title: "penguins"
output: pdf_document
date: "2025-11-18"
---

# Notes Example

```{r}
data("penguins")


dream_pen <- subset(penguins, island == 'Dream')
tor_pen <- subset(penguins, island == 'Torgersen')

hist(dream_pen$flipper_len)
hist(tor_pen$flipper_len)

t.test(dream_pen$flipper_len,
       tor_pen$flipper_len,
       alternative = 'two.sided',
       mu = 0,
       conf.level = .95)

```


# Example to try on your own

We are ignoring tension for the moment if you are wondering. 

```{r}
data("warpbreaks")

yarn_a <- subset(warpbreaks, wool == 'A')
yarn_b <- subset(warpbreaks, wool == 'B')

```

1) State you null and alternative hypothesis

2) Visualize your data


```{r}
library(ggplot2)

ggplot(warpbreaks,
       aes(breaks)) + 
  geom_histogram() +
  facet_grid(~wool) +
  theme_bw()
```

3) Check your assumptions (random, independent and identically distributed)

4) Regardless if we fail our assumptions, let's continue. Using R, calculate the test-statistic and the p-value

5) Give a decision for your test

6) State and interpret a 95\% confidence interval for the difference in their means