---
title: "Code_Along_115"
output: html_document
date: "2025-09-22"
editor_options: 
  chunk_output_type: console
---

This is to help 
1. Reenforce the coding that we have gone over and 
2. Help make a boring lecture slightly less boring


```{r}
colleges <- read.csv("https://remiller1450.github.io/data/Colleges2019_Complete.csv")

head(colleges) #look at the first few rows of data

colnames(colleges) #and look at the column names
```


```{r}
library(ggplot2) #load ggplot2

ggplot(data = colleges, #data is colleges
       aes(x = PercentFemale)) + #and our x-axis is % female
  geom_histogram() + #and I want a histogram
  theme_bw() #with a black and white theme


#summary summarizes information  about the thing fed to it
summary(colleges$PercentFemale)
#in particular the five number summary (plus the mean) 
#for a numeric vector
```



```{r}
#same battleplan as the above ones but now we switch
#out PercentFemale with Enrollment
ggplot(data = colleges,
       aes(x = Enrollment)) +
  geom_histogram() +
  theme_bw()

#and this time I swithced out geom_histogram for geom_boxplot
ggplot(data = colleges,
       aes(x = Enrollment)) +
  geom_boxplot() +
  theme_bw()

#summary summarizes information  about the thing fed to it
summary(colleges$Enrollment)
```



# Enrollment on Log base 10 Scale
```{r}
#Very little has to change for the graph!
ggplot(data = colleges,
       aes(x = Enrollment)) +
  geom_histogram() +
  theme_bw() +
  scale_x_log10() #this is all we add, it says the x-axis is on
                  #a log base 10 scale


#################
#Calculating the five number summary is a bit harder
#We have to make a new variable via a transformation 


#new variable time
#We are creating a new column on our data set
#called enrollment_log10. We are then assigning
#the right hand side to that column.

#The right hand side is the function log() applied
#to our enrollment buuut we have changed the base 
#of the log to base = 10 (instead of e)
colleges$enrollment_log10 <- log(colleges$Enrollment,
                                 base = 10)

colnames(colleges) #glance at the column names to see
#if our new variable is there

#and now we look at the summary
summary(colleges$enrollment_log10)
```


# Correcting the Wrong Axis
Shout out to your mentor for figuring out which variable was used to make the graph. It's FourYearComp_Females it turns out. Hard to find the right variable but the code to make the graph isn't hard at all
```{r}
ggplot(data = colleges,
       aes(x = FourYearComp_Females)) +
  geom_histogram() +
  theme_classic() 
```